Constraining gravitational wave amplitude birefringence with GWTC-3

### 2023-03-06 src/tex/ms.tex
Add GW190521 freq independent plot

The fact that the birefringent analysis favors a high $\kappa$ rather than a high $\chi_p$ is a consequence of prior volume: many more configurations are available with long distances and high $\kappa$ than with short distances and small $|\kappa|$.
The preference for $\kappa > 0$ over $\kappa < 0$ (and, therefore, the lack of bimodality in the $\cos\iota$ and $\kappa$ posteriors), is likely related to both the definite measurement of left-handed polarizations and the specific phasing of the precession cycle.
The latter manifests as a precise constraint on the spin orientation and phase angles in the \ac{GR} analysis \cite{Varma:2021csh}.
The observed amplitude modulation (say, increasing vs decreasing towards the merger) likely determines the allowed sign of $\kappa$ for this event.
Indeed, the inferred polarization angle in the birefringent case is notably different from the \ac{GR} one.
We provide the full corner plot for all relevant parameters in Fig.~\ref{fig:corner_GW170818_appendix} in Appendix XX.

@@

\begin{figure}[h]
    \script{corner_GW190521.py}
    \includegraphics[width=\columnwidth]{figures/corner_GW190521.pdf}
    \caption{
        GW190521 posterior on $\kappa$, luminosity distance $d_L$ and inclination $\cos\iota$ from our birefringence analysis (blue), compared to the GR result (orange).
        % The posterior of $\kappa$, luminosity distance $d_L$ and $\cos{\iota}$ for GW190521.
        % Colors in the plot represent the \ac{PE} result with \ac{GR} and the frequency-dependent birefringence, respectively.
        % The 2D contours correspond to the $39.35\%$ ($1\sigma$) and $90\%$ credible levels.
        % This plot shows that frequency-dependent birefringence cannot break the degeneracy between $\kappa$ and $\iota$.
    }
    \label{fig:corner_GW190521}
\end{figure}

@@

\section{GW170818}

>>>>

The fact that the birefringent analysis favors a high $\kappa$ rather than a high $\chi_p$ is a consequence of prior volume: many more configurations are available with long distances and high $\kappa$ than with short distances and small $|\kappa|$.
The preference for $\kappa > 0$ over $\kappa < 0$ (and, therefore, the lack of bimodality in the $\cos\iota$ and $\kappa$ posteriors), is likely related to both the definite measurement of left-handed polarizations and the specific phasing of the precession cycle.
The latter manifests as a precise constraint on the spin orientation and phase angles in the \ac{GR} analysis \cite{Varma:2021csh}.
The observed amplitude modulation (say, increasing vs decreasing towards the merger) likely determines the allowed sign of $\kappa$ for this event.
Indeed, the inferred polarization angle in the birefringent case is notably different from the \ac{GR} one.
We provide the full corner plot for all relevant parameters in Fig.~\ref{fig:corner_GW170818_appendix} in Appendix~\ref{sec:corner_GW170818_appendix}.

@@

\begin{figure}[h]
    \script{corner_GW190521.py}
    \includegraphics[width=\columnwidth]{figures/corner_GW190521.pdf}
    \caption{
        GW190521 posterior on $\kappa$, luminosity distance $d_L$ and inclination $\cos\iota$ from our frequency-dependent (blue) and frequency-independent (orange) birefringence analysis, compared to the GR result (green).
        % The posterior of $\kappa$, luminosity distance $d_L$ and $\cos{\iota}$ for GW190521.
        % Colors in the plot represent the \ac{PE} result with \ac{GR} and the frequency-dependent birefringence, respectively.
        % The 2D contours correspond to the $39.35\%$ ($1\sigma$) and $90\%$ credible levels.
        % This plot shows that frequency-dependent birefringence cannot break the degeneracy between $\kappa$ and $\iota$.
    }
    \label{fig:corner_GW190521}
\end{figure}

@@

\section{Full corner plot of GW170818}
\label{sec:corner_GW170818_appendix}
